---
title: "Using `axamodel` package "
author: "Etienne Kintzler"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc : true 

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = F}
knitr::opts_chunk$set(cache = TRUE, message = FALSE)
```

This package aims at creating model for *Generalized Linear Models* (GLM). More specifically, it includes variable selection and interactions detection. The variable selection is performs with *Least Absolute Shrinkage and Selection Operator* (LASSO) applied to GLM. The interactions are detected with *Gradient Tree Boosting* (GTB) and Friedman’s *H-statisti*c.

In this vignette the two principal functions of the package are presened : 

* the `lassoSelection` function used to select variables

* the `detectInteractions` function used to detect interactions and compute the resulting gain in GLM modeling. 

## Installation

The package can be installed via git hub using :


```{r, eval = F}
devtools::install_git('https://github.com/etiennekintzler/automaticmodel.git')
```

or

```{r, eval = F}
devtools::install_git('https://github.axa.com/ds4a/axamodel.git')
```

Also, this package use the development version of the package gbm which can be acquired using :

```{r, eval = F}
devtools::install_git('https://github.com/gbm-developers/gbm')
```




```{r, include = T, echo = F, results = 'hide', message = F}
library(axamodel)
library(dplyr)
```

### Loading the dataset

We will use the dataset *portfolio_example* in the package axaml

```{r}
data <- axaml::portfolio_example
df   <- subset(x = data, select = -c(CalYear, Indtppd, Indtpbi, Numtppd, PolNum)) %>% 
  cleanData(n.levels = 10)
```

The `cleanData` function will clean the dataset removing variable with too many NA, too many levels, or too much correlation. The target variable is the number of claims `Numtpbi`, and takes its value in $\{0,1,2,4\}$. It will be thus modelled using poisson family.


## 1.  Using `lassoSelection` to select variables

### 1.1. Short introduction 
The `lassoSelection` returns the best linear model using LASSO-GLM procedure. Lasso is $L_1$-penalized least squares and performs variable selection through shrinking some estimates to zero. The program solved is the following :
\[
\min_\beta \left\lVert \mathbf{y}- \mathbf{X}\beta \right\rVert_2^2 + \left\lVert \beta\right\rVert_0
\]
The value of $\lambda$ is determined using k-fold cross-validation. The criterion used for the cross-validation is passed 
to `lassoSelection` function with parameter `type.measure`.

## 1.2. Function call

The call for the `lassoSelection`is the following :

```{r, eval = T, echo = T, cache = T}
best.glm <- lassoSelection(data           = df %>% select(-Exposure), 
                           type.measure   = 'deviance',
                           family         = 'poisson', 
                           lambda.type    = 'lambda.min',
                           target         = 'Numtpbi',
                           offset         = log(df$Exposure),
                           nlambda        = 10,
                           nfold          = 5)
```


The function is based on `glmnet::cv.glmnet` for all the distributions but for gamma. For gamma distribution
`HDtweedie::cv.HDtweedie` is used. 

It is possible to add an offset to the function except for the gamma family (since `HDtweedie::cv.HDtweedie` does not enable it). If the offset is provided it is necessary to remove it from the data passed to the function.

### 1.2. Output diagnostic
The output of `lassoSelection` will be on object of class `LassoGLM` which elements can be accessed and plot via differents methods. Four methods will be presented here:

* `perf` which returns the performances (Root Mean Squared Error and Gini index) of the best model according to the Lasso-GLM procedure.

* `plotLambda` which plots the value of type.measure according to the different values of $\lambda$

* `plot` which plots the Lorenz curve of the best model with the resulting Gini Index

* `summary` provides a table with the selected variable and the non-selected variables.


```{r}
plotLambda(best.glm)
```

The value of lambda that optimize (minimize or maximize) `type.measure`, called `lambda.min`, is around $\exp(-6)$.
The lambda that gives the most regularized model such that `type.measure` is within one standard error of its optimum, called `lambda.1se` is about $\exp(-4)$


```{r}
perf(best.glm)
```

The following method plot the Lorenz Curve and the resulting Gini index.

```{r}
plot(best.glm)
```

The following command returns the variable selected by the procedure

```{r}
summary(best.glm)
```

One can also access the formula of this *base model* using :

```{r}
best.glm$formula
```



## 2. Detecting interactions

The detection of interactions is performed by `detectInteractions` function.

### 2.1. Short introduction

An interaction between two variables is a form of *non-additivity* of the response function. Let $F(x_1, x_2)$ the response function, there is *no* interaction between $x_1$ and $x_2$ if the response function writes :
$$F(x_1, x_2) = f_1(x_1)+ f_2(x_2) $$
where $f_i$ is a function that only depends on the variable $i$.

The *H-statistic* is a quantitty used within GBT framework to assess the presence of an interaction.
Considerong, two variables $x_1$ and $x_2$, the quantity writes:

\[
H_{1,2} = \frac{\sum(\widehat{F}_{1,2}(x_1, x_2)-\widehat{F}_1(x_1) - \widehat{F}_2(x_2)  )^2}{\sum(\widehat{F}_{1,2}(x_1, x_2))^2}
\]

Where $F_s, s\subset \{1,..,n\}$ is the partial dependence function (see Friedman J.H. & Popescu, B. E. (2008). Predictive Learning via Rule Ensembles. *The Annals of Applied Statistics*, 916-954)

### 2.2. Function call

```{r, eval = T, echo = T, cache = T, results = "hide"}
gbm.control <- list(distribution         = 'poisson', 
                    n.trees              = 3, 
                    interaction.depth    = 5, 
                    shrinkage            = .1, 
                    bag.fraction         = .5,
                    train.fraction       = .75, 
                    importance.threshold = 0,
                    max.select           = 20,
                    n.sample             = 1e3)

glm.control <- list(family           = poisson(link = 'log'), 
                    train.fraction   = .75, 
                    include          = F,
                    threshold.values = seq(0, 1, length.out = 20),
                    seed             = 1)

interactions <- detectInteractions(data             = df %>% select(-Exposure), 
                                   target           = 'Numtpbi',
                                   offset           = df$Exposure,
                                   formula.best     = best.glm$formula,
                                   max.interactions = 30, 
                                   gbm.control      = gbm.control,
                                   glm.control      = glm.control)
```

The *base model* is specified in formula form in the parameter *formula.best*. The formula can either but the output of `lassoSelection` or can be user defined. This *base model* is compared with the *augmented model* that includes interactions(s).


The `gbm.control` argument takes the form of a list and contains the parameters necessary for fitting a gbm using `gbm::gbm` such as `train.fraction`, `shrinkage`, `bag.fraction`, `n.trees` and `distribution` and `interaction.depth`.The other parameters in the list such as `importance.threshold`, `n.sample`, `max.select` are related to the computation of the *H-statistic*.

The `glm.control` argument also takes the form of a list and contains parameters used to fit a glm using `stats::glm`. The `threshold.values` is the threshold for the *H-statistic* according to which the interactions are selected. There are as much *augmented models* as elements in `threshold.values`. The **best** `augmented model` is the one that maximise the performance criterion (Gini index or RMSE).

### 2.3. Output diagnostic

Four methods can be used to diagnostic the output of detectInteractions :

* `perf` that compares the performance of the base model vs the augmented model.

* `summary` give the interactions with the highest H-statistic value.

* `plotBest` plot the Lorenz curve of both the base model and the augmented model along with the gini gain.

* `plot` is the interaction plot of the i-th best interaction

* `returnBest` return the best *augmented model*.


```{r}
perf(interactions)
```

There is no substantial differences between the best *augmented model* and the *base model*.

```{r}
summary(interactions)
```

These are the interactions included in the best *augmented model*.

```{r}
plot(interactions, simplify = 8)
```

This is the interaction plot for the best interaction found
(the best is the one with the highest *H-statistic*)

